An assessment of photometric redshift PDF performance in the context of LSST

### 2018-11-13 paper/tex/results.tex
address comments for section 5.2

As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $\approx 200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of roughly $\delta z = 0.01$ for each code.
A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.
In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others.

@@

Fig.~\ref{fig:pitqq} shows both the quantile-quantile plots (red) and the histogram of PIT values (blue) summarizing the results from each photo-$z$ code.  The red line shows the measured quantiles, while the black diagonal represents the ideal QQ values if the distribution were perfectly reproduced.  A second panel below the main panel for each code shows the difference between $Q_{data}$ and $Q_{theory}$, i.~e.~the departure from the diagonal, for clarity.
Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.  An overall bias where the predicted redshift is systematically low manifests as the measured QQ value falling above the diagonal, as is the case for \textsc{BPZ} and \textsc{EAZY}, while a systematic overprediction shows up as the measured QQ value falling below the diagonal, as seen in \textsc{TPZ}.  In terms of PIT histograms, a systematic underprediction of redshift corresponds to fewer PIT values at $PIT<0.5$ and more at $PIT>0.5$, while a systematic overprediction will show the opposite.

Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with an dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.  \textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.  In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.  The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.  for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.  A similar sharpening procedure could be beneficial for several codes.
Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.

@@

%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{fig/PIT_histogram_plot_11codes.jpg}
%\caption{The PIT histogram plots produced for every photo-$z$ code.  Ideal behaviour would match a uniform distribution, indicated by the horizontal dashed line.  Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.  Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.  Asymmetric features are indicative of systematic bias in the redshift predictions.
%\aim{Could we combine this with the previous figure by plotting both as differences from ideal on one plot with left and right axis labels?  The information seems degenerate (which makes sense!)}\red{as you say, the information is degenerat, this form emphasizes the differences a bit more, and shows off the catastrophic outliers with PIT=0 and 1 a bit better in my mind.} } \label{fig:pit}
%\end{figure*}
\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.  The statistics are often highly correlated, though the AD statistic truncates the extrema of the distribution and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
\end{figure*}

@@

%end of new text (and slightly modified text)
Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks with the largest number of training galaxies, and missing probability in the troughs where training galaxies are of fewer number.
Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations of \textsc{ANNz2} and \textsc{METAPhoR}.
A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining for these particular codes.
%\aim{Move the text on Scott's Rule to Discussion?}
%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all twelve codes.
%As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
%While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.
\textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.  \textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.  Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.  Either of these decisions could be the cause of the bias seen in our results.  Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.

@@

Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution, including the moments of the ``truth'' distribution for comparison.  Several codes are able to reproduce the mean and variance of the distribution to less than a per cent, while several codes do not, which may be a cause for concern, given that mean and variance of the redshift distribution are key properties in cosmological analyses.  We note that this stated goal of the study as defined for participants was to accurately reproduce $p(z)$, the ``stacking'' of the probability distributions to estimate $\hat{N}(z)$ was not the focus as stated to the participants.  This explains why some of the best-performing empirical codes in terms of $p(z)$ measures (e.~g.~\textsc{FlexZBoost}) do not do as well at reproducing $\hat{N}(z)$ moments.  Had we defined a different parameter to optimize, in this case overall accuracy of $\hat{N}(z)$ rather than individual $p(z)$, would result in improved performance in a particular metric.  That is, optimizing photo-$z$ performance for one metric does not automatically give optimal performance for other metrics.  As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.  In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.  We discuss more realistic tomographic bin selections that will be explored in a follow-up paper in Section~\ref{sec:futurework}.
%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.  Why is this?  Is it the over/underprediction of the high redshift part of the distribution?  Discuss this after talking with Rafael.}

>>>>

As stated in Section~\ref{sec:metrics}, $p(z)$ is parameterized as $200$ piecewise constant bins covering $0<z<2$ for all twelve codes, giving a grid size of $\delta z = 0.01$ for each code.
A piecewise constant grid was a natural choice for some photo-$z$ codes, for instance most template-based codes compute likelihoods on a fixed grid.  In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others in a limited number of cases, for example bright galaxies with very narrow $p(z)$ where the true peak falls between grid points.  We will discuss PDF storage in Section~\ref{sec:discussion}.

@@

Fig.~\ref{fig:pitqq} shows both the quantile-quantile plots (red) and the histogram of PIT values (blue) summarizing the results from each photo-$z$ code.  The red line shows the measured quantiles, while the black diagonal represents the ideal QQ values if the distribution were perfectly reproduced.  A second panel below the main panel for each code shows the difference between $Q_{\rm data}$ and $Q_{\rm theory}$, i.~e.~the departure from the diagonal, for clarity.
Biases and trends in whether the average width of the $p(z)$ values being over/under-predicted are evident.  An overall bias where the predicted redshift is systematically low manifests as the measured QQ value falling above the diagonal, as is the case for \textsc{BPZ} and \textsc{EAZY}, while a systematic overprediction shows up as the measured QQ value falling below the diagonal, as seen in \textsc{TPZ}.  In terms of PIT histograms, a systematic underprediction of redshift corresponds to fewer PIT values at $PIT<0.5$ and more at $PIT>0.5$, while a systematic overprediction will show the opposite.

Examination of the PIT histograms and QQ plots shows that there are fairly generic issues with the width of $p(z)$ uncertainties: \textsc{Delight, CMNN, SkyNet} and \textsc{TPZ} all show a PIT histogram with a dearth of low values and and an excess of high values, signs that, on average, their $p(z)$ are more broad than the true distribution of redshifts.  \textsc{METAPhoR} shows the opposite trend, indicating the the $p(z)$  are more narrow than the distributions given by the true redshifts.  In all of these code cases there is a free parameter or bandwidth that can be used to tune uncertainties.  The sensitivity of multiple codes to this bandwidth choice emphasizes the fact that great care must be taken in setting user-defined parameters in photo-$z$ codes, even in the presence of representative training/validation data.  for \textsc{FlexZBoost} the ``sharpening'' parameter (described in Section~\ref{sec:flexzboost}) plays a key role in improving the results, resulting in a QQ plot that is very nearly diagonal.  A similar sharpening procedure could be beneficial for several codes.
Interestingly, the three purely template-based codes, \textsc{BPZ, EAZY}, and \textsc{LePhare}, show relatively well behaved $p(z)$ statistics (albeit with some bias), which may indicate that the likelihood estimation with representative templates is accurately capturing the uncertainties on individual redshifts.

@@

%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{fig/PIT_histogram_plot_11codes.jpg}
%\caption{The PIT histogram plots produced for every photo-$z$ code.  Ideal behaviour would match a uniform distribution, indicated by the horizontal dashed line.  Histograms with an overabundance of PIT values at the centre of the distribution indicate $p(z)$ distributions that are overly broad, while an excess of values at the extrema indicate $p(z)$ distributions that are overly narrow.  Values of PIT=0 and PIT=1 indicate ``catastrophic failures'' where the true redshift is completely outside the support of $p(z)$.  Asymmetric features are indicative of systematic bias in the redshift predictions.
%\aim{Could we combine this with the previous figure by plotting both as differences from ideal on one plot with left and right axis labels?  The information seems degenerate (which makes sense!)}\red{as you say, the information is degenerat, this form emphasizes the differences a bit more, and shows off the catastrophic outliers with PIT=0 and 1 a bit better in my mind.} } \label{fig:pit}
%\end{figure*}
\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig/KSvsCvMvsAD_PIT_withnull_jpg.jpg}
\caption{A visual representation of the Kolmogorov-Smirnoff (KS, blue diamond), Cramer-von Mises (CvM, black star), and Anderson-Darling (AD, red asterisk) statistics for the PIT distributions.  The statistics are often highly correlated, with similar relative values between metrics for each code; however, the AD statistic excludes some values at the extrema of the distributions, and can have disparate values compared to KS and CvM.} \label{fig:pit_stats}
\end{figure*}

@@

%end of new text (and slightly modified text)
Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks with the largest number of training galaxies, and missing probability in the troughs where training galaxies are of fewer number.
Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations and bandwidth choices of \textsc{ANNz2} and \textsc{METAPhoR}.
This, once again, emphasizes that care much be taken in choice of bandwith parameters in individual codes in order to avoid such overtraining.
%\aim{Move the text on Scott's Rule to Discussion?}
%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all twelve codes.
%As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
%While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.
\textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.  \textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.  Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.  Either of these decisions could be the cause of the bias seen in our results.  Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.

@@

Table~\ref{tab:moments} lists the first three moments of the stacked $\hat{N}(z)$ distribution, including the moments of the ``truth'' distribution for comparison.  Several codes are able to reproduce the mean and variance of the distribution to less than a per cent, while several codes do not, which may be a cause for concern, given that mean and variance of the redshift distribution are key properties in cosmological analyses.  We note that this stated goal of the study as defined for participants was to accurately reproduce $p(z)$, the ``stacking'' of the probability distributions to estimate $\hat{N}(z)$ was not the focus as stated to the participants.  This explains why some of the best-performing empirical codes in terms of $p(z)$ measures (e.~g.~\textsc{FlexZBoost}) do not do as well at reproducing $\hat{N}(z)$ moments.  Had we defined a different parameter to optimize, in this case overall accuracy of $\hat{N}(z)$ rather than individual $p(z)$, would result in improved performance in a particular metric.  That is, optimizing photo-$z$ performance for one metric does not automatically give optimal performance for other metrics.  As previously stated, there are a variety of scientific use cases for photo-$z$'s in large upcoming surveys, and care must be taken in how the metrics used to optimize catalog photometric redshifts are defined as well as in how they are used.  This implies that we may need multiple photo-z estimators, tuned to the particular metric, in order to maximize science returns in upcoming surveys.  In addition, very few scientific use cases will employ the overall $\hat{N}(z)$ with no cuts, as we explore in this paper.  We discuss more realistic tomographic bin selections that will be explored in a follow-up paper in Section~\ref{sec:futurework}.
%\red{[mention that we are calculating moments for the entire sample, will look at fiducial tomographic bins in a follow-up paper (unless group and reviewers realy think we should include in this paper.]}
%\red{FlexZBoost has some of the best $n(z)$ statistics, but the moments are not good.  Why is this?  Is it the over/underprediction of the high redshift part of the distribution?  Discuss this after talking with Rafael.}
